---
title: "Data Annotation"
author: "Interdisciplinary Team"
date: "January 27, 2016"
output: html_document
---

```{r echo=FALSE}
library(ggplot2)
library(reshape2)
```

Some summaries of the hand classification:
```{r hand_classified}
good_runs <- data.frame(subset(all_runs, msg < 55778),
                        variable = NA)

# Anything that indicates a job:
class_week <- subset(melt(good_runs, measure.vars = c(3,6,7), id.vars = 1), value == TRUE)

class_week <- data.frame(class_week, apply(weeks, 2, as.numeric)[class_week$msg,])
class_week$date <- class_week$V2/12 + class_week$V3

multi_class <- which(rowSums(table(class_week[,1:2])>0)>1)

hand_weeks <- apply(weeks[good_runs$msg,], 2, as.numeric)


ggplot(class_week) + geom_bar(aes(x = date, fill = variable)) + 
  facet_wrap(~variable, ncol = 1) + 
  coord_cartesian(xlim=c(2000, 2015), expand = FALSE) +
  theme_minimal() +
  theme(legend.position = 'none')

```

The number of messages hand classified by month, by class (gp - Graduate; tt - tennure track)


Out of the `r nrow(all_runs)` messages, there are `r length(multi_class)` messages that have been classified into more than one job class.  In every case, except one it is a position that is classed as both "Graduate" and "Postdoc".  In the other case it's a tenure-track/postdoc posting.

Most of these are joint postings looking for either a postdoc or a high quality doctoral student.

```{r classified_by_class, echo = FALSE}

all_dat = readRDS('data/all_melt.RDS')

ggplot(all_dat) + 
  geom_bar(aes(x = date, y = value), stat='identity') +
  facet_wrap(~class, ncol = 1) +
  theme_minimal()

```
**Figure 1**. Posting volume for job adds in each class.


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
